---
title: 'Chapter 6: 6.6, 6.12, 6.20, 6.28, 6.44, 6.48'
author: "Max Wagner"
date: "October 31, 2015"
output:
  pdf_document: default
  html_document:
    highlight: espresso
    theme: flatly
---
***

###6.6
####a. 
False, it should be about the population, not the sample.

####b.
True, a fixed version of (a).

####c.
Weird wording(false), but the point is 95% of the samples should include the population proportion.

####d.
False, the margin of error should be lower/narrower with a lower CI.

###6.12
####a.
48% represents a sample stat.

####b.
```{r}
n <- 1259
legal <- .48
se <- sqrt((legal * (1 - legal)) / n)
z <- qnorm(.975)
me <- z * se
```
The CI is `r legal` +/- `r me`.

####c.
Being that the percentages are 52% and 48% (almost even), it is likely close to normal.

####d.
Technically the interval for legality can reach up to `r legal + me`. It is a stretch but not technically wrong.

###6.20
```{r}
me <- .02
z <- qnorm(.975)
se <- me / z
n <- (legal * (1 - legal)) / se ^ 2;n
```

###6.28
```{r}
ca <- .08
ca_n <- 11545
or <- .088
or_n <- 4691
diff <- or - ca
se <- sqrt(((ca * (1 - ca)) / ca_n) + ((or * (1 - or)) / or_n))
z <- qnorm(.975)
me <- z * se
```
The CI is `r diff` +/- `r me`. This interval contains 0, so they are statistically different.

###6.44
####a.
Null: Deer do not prefer one habitat over another, and foraging is proportional to land distribution, Alt: Deer do prefer certain habitats, and foraging is not proportional to land distribution.

####b.
One way is to use a chi-square test.

####c.
The conditions state the sample should be independent, and expecteded counts should be 5 or greater.

####d.
```{r}
expected <- c(.048*426, .147*426, .396*426, (1 - .396 - .147 - .048)*426)
actual <- c(4, 16, 67, 345)
df <- length(actual) - 1
chi <- sum(((actual - expected) ^ 2) / expected)
p <- pchisq(chi, df, lower.tail = FALSE)
```

P is almost 0, which means we reject the null, and say that deer do prefer to forage in someplaces over others.

###6.48
####a. 
yay more chi square

####b.
Null: depression and coffee consumption are unrelated, Alt: depression and coffee consumption are related.

####c.
do: `r 2607 / 50739`, don't: `r 48132 / 50739`

####d.
```{r}
exp <- 2607 * (6617/50739);exp
contrib <- (373 - exp) ^ 2 / exp;contrib
```

####e.
```{r}
pchisq(20.93, df = 4, lower.tail = FALSE)
```

####f.
The p value is less than .05, so we reject the null, and accept that there is a relationship between coffee and depression.

####g.
I agree with the author for a few reasons. The first being that a p value of .03 is convincing, but not conclusive. The second being that there could a whole lot of other variables going on these women's lives.