---
title: 'Bayesian: 2.1, 5.1, 5.2, '
author: "Max Wagner"
date: "Decemeber 10, 2015"
output:
  pdf_document: default
  html_document:
    highlight: espresso
    theme: flatly
---
***

###2.1

A. Model A has no bias, as it is an even 1/4 chance for each side. Which gives a P(x) = 1/4 for each side.

B. Model B gives a higher chance as the number increases. P(x) = `r c(1:4) / 10`

C. Model C gives a higher chance as the number decreases. P(x) = `r 12 / (25 * c(1:4))`

###5.1

Prob for first test, and then the second...
```{r}
present <- 0.001
absent <- 1 - present
present_pos <- 0.99
absent_pos <- 0.05

top <- present_pos * present
bottom <- present_pos * present + (absent_pos * (1 - present))
first_test <- top / bottom;first_test

top <- (1 - present_pos) * first_test
bottom <- (1 - present_pos) * first_test + ((1 - absent_pos) * (1 - first_test))
second_test <- top / bottom;second_test
```

###5.2
####A.

The table from part can be seen below:
```{r}
library(knitr)
sad <- c(0.99 * 0.001 * 100000, (1.0 - 0.99) * 0.001 * 100000, 0.001 * 100000)
happy <- c(0.05 * (1.0 - 0.001) * 100000, (1.0 - 0.05) * (1.0 - 0.001) * 100000, (1.0 - 0.001) * 100000)
tot <- sad + happy
sht <- cbind(sad, happy, tot)
rownames(sht) <- c("+", "-", "tot")
kable(sht)
```

####B.

From the table, it would appear that every 99 / 5,094 would have the disease. The actual number is the same as the percentage from above, 99 / 5094 = `r 99 / 5094`.

####C.

The tree is represented as a table.
```{r}
left <- c(10000, 10000 * 0.99, 9900 * (1.0 - 0.99))
right <- c(9990000, 9990000 * 0.05, 499500 * (1.0 - 0.05))
lr <- cbind(left, right)
kable(lr)
```

####D.

The proportion works out to be the same.
```{r}
99 / (99 + 474525)
```
